% ConvergenceSafetySpec.tex — LaTeX source of truth for the convergence-safety formal specification
% Bootstrapped from the formal-apparatus at Phase 3 (LaTeX elicitation)
% Co-evolves with theories/ConvergenceSpec.v pre-lock; LaTeX is canonical post-lock.
% Status: LOCKED v0.2 (2026-06-11) — matches coqchk-passing ConvergenceSpec.v (16 theorems)

\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{cleveref}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{notation}{Notation}[section]
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

% Project-specific macros
\newcommand{\Converged}{\mathsf{Converged}}
\newcommand{\Diverged}{\mathsf{Diverged}}
\newcommand{\chk}[2]{\mathtt{check}(#1,\,#2)}
\newcommand{\isvarying}[1]{\mathtt{is\_varying}(#1)}
\newcommand{\barrierfree}[1]{\mathtt{barrier\_free}(#1)}
\newcommand{\cdcf}[1]{\mathtt{has\_diverging\_cf}(#1)}
\newcommand{\mergedim}{\mathbin{\sqcup}}
\newcommand{\emptydim}{\bot_d}
\newcommand{\incl}{\subseteq}

\title{Convergence Safety --- A Formal Specification\\
  \large \textit{Sarek GPU barrier safety analysis}}
\author{Sarek formal apparatus}
\date{2026-06-11 (v0.2 LOCKED)}

\begin{document}

\maketitle

\begin{abstract}
We give a formal specification of the convergence-safety analysis pass in
\texttt{Sarek\_convergence.ml}, the GPU barrier-safety checker for the Sarek
OCaml GPU DSL. The analysis tracks an \emph{execution mode} ($\Converged$ or
$\Diverged$) through an abstract expression type and collects barrier errors
when a synchronisation point is reached inside diverged control flow.
We prove sixteen properties: five join-semilattice laws for the dimension-usage
monoid, one homomorphism law for sequential composition, six T1 safety
theorems about the mode-tracking state machine (including the F-01
\texttt{ESuperstep} outer-diverged error property), and four T1 theorems
for the env-threaded model introduced in T2-F02 to handle let-alias
variability propagation.
\end{abstract}

%───────────────────────────────────────────────────────────────────────────────
\section{Overview}
\label{sec:overview}

\subsection*{Target component}

\texttt{Sarek\_convergence.ml} in \texttt{sarek/ppx/}. Functions of interest:
\begin{itemize}
  \item \texttt{is\_thread\_varying : texpr -> bool} --- conservative check for
        thread-varying subexpressions.
  \item \texttt{check\_expr : convergence\_ctx -> texpr -> error list} ---
        barrier-safety error collector.
  \item \texttt{contains\_diverging\_control\_flow : texpr -> bool} ---
        superstep analysis predicate.
\end{itemize}

\subsection*{Properties claimed}

\begin{enumerate}
  \item[\textbf{T0}] Dimension-usage forms a join-semilattice (\cref{sec:dimmerge}).
  \item[\textbf{T0}] Sequential composition of checks is homomorphic (\cref{thm:seqhom}).
  \item[\textbf{T1}] \emph{Diverged-clean iff barrier-free} (\cref{thm:div-clean}).
  \item[\textbf{T1}] Mode strengthening is monotone (\cref{thm:monotone}).
  \item[\textbf{T1}] Non-varying-condition implies clean check (\cref{thm:cdcf-clean}).
  \item[\textbf{T1}] Varying condition on non-barrier-free body always flags an error
        (\cref{thm:varying-flags}).
  \item[\textbf{T1}] Outer-diverged \texttt{ESuperstep} always raises a barrier error
        (F-01, \cref{thm:superstep-outer-diverged}).
  \item[\textbf{T1}] Env-lookup after env-extend returns the stored value
        (\cref{thm:env-lookup-extend-same}).
  \item[\textbf{T1}] Let-alias variability is propagated through \texttt{ELet}
        (\cref{thm:env-let-alias-varying}).
  \item[\textbf{T1}] \texttt{EVar} carries no barrier under $\Diverged$ mode
        (\cref{thm:env-var-diverged-clean}).
  \item[\textbf{T1}] \texttt{check\_env} catches barriers reachable via let-alias
        (F-02, \cref{thm:env-check-let-alias-catches}).
\end{enumerate}

\subsection*{Out of scope (T2/T3)}

Semantic soundness of \texttt{is\_varying} (needs $\texttt{texpr}$ execution
semantics), deadlock freedom (needs lockstep execution model),
and warp divergence errors.

%───────────────────────────────────────────────────────────────────────────────
\section{Notation and types}
\label{sec:notation}

\begin{definition}[Execution mode]
$\mathsf{ExecMode} ::= \Converged \mid \Diverged$
\end{definition}

\begin{definition}[Error]
$\mathsf{Error} ::= \mathsf{BarrierError}$
\end{definition}

\begin{definition}[Abstract expression]
\label{def:expr}
\[
  e ::=
    \mathsf{ELit} \mid \mathsf{EVary} \mid \mathsf{EBarrier}
  \mid \mathsf{EVar}(x)
  \mid \mathsf{EBinop}(e_1, e_2)
  \mid \mathsf{EUnop}(e)
  \mid \mathsf{EIf}(e_c, e_t, e_f)
  \mid \mathsf{EWhile}(e_c, e_b)
  \mid \mathsf{EFor}(e_{lo}, e_{hi}, e_b)
  \mid \mathsf{ESeq}(\bar{e})
  \mid \mathsf{ELet}(x, e_v, e_b)
  \mid \mathsf{ESuperstep}(b, e_{body}, e_{cont})
  \mid \mathsf{EApp}(\bar{e})
\]
where $\bar{e}$ ranges over $\mathsf{expr}^*$, $x \in \mathbb{N}$ is a
variable identifier, and $b \in \mathbb{B}$ is a Boolean
\emph{divergent flag} (\textsf{true} = outer superstep is already divergent).
\end{definition}

\begin{notation}
$\isvarying{e}$ denotes the Boolean predicate that returns $\mathsf{true}$
when $e$ contains a thread-varying leaf.
$\barrierfree{e}$ denotes the Boolean predicate that returns $\mathsf{true}$
when $e$ contains no $\mathsf{EBarrier}$ leaf.
$\cdcf{e}$ denotes the Boolean predicate that returns $\mathsf{true}$
when $e$ contains a diverging control-flow sub-expression.
$\chk{m}{e}$ denotes the list of errors emitted when checking $e$ in
execution mode $m$.
\end{notation}

\begin{definition}[Dimension-usage record]
\label{def:dimuse}
\[
  d ::= \{\, \mathsf{uses\_x},\, \mathsf{uses\_y},\, \mathsf{uses\_z},\,
            \mathsf{uses\_block\_dim},\, \mathsf{uses\_grid\_dim},\,
            \mathsf{uses\_thread\_idx},\, \mathsf{uses\_block\_idx},\,
            \mathsf{uses\_shared\_mem} \,\}
\]
Each field is a Boolean.
$\emptydim$ denotes the all-false record.
$\mergedim$ denotes field-wise disjunction.
\end{definition}

%───────────────────────────────────────────────────────────────────────────────
\section{Semantic rules}
\label{sec:semantics}

\subsection{$\isvarying{e}$}

This is the \emph{binding-blind} variant: $\mathsf{EVar}(x)$ is treated as
non-varying regardless of what was bound to $x$.  The env-threaded variant
$\mathtt{is\_varying\_in\_env}$ (\cref{sec:envthreaded}) corrects this.

\[
  \isvarying{\mathsf{EVary}} = \top
  \quad
  \isvarying{\mathsf{ELit}} = \bot
  \quad
  \isvarying{\mathsf{EBarrier}} = \bot
  \quad
  \isvarying{\mathsf{EVar}(x)} = \bot
\]
\[
  \isvarying{\mathsf{EBinop}(a,b)} = \isvarying{a} \lor \isvarying{b}
  \qquad
  \isvarying{\mathsf{EUnop}(e)} = \isvarying{e}
\]
\[
  \isvarying{\mathsf{EIf}(c,t,f)} = \isvarying{c} \lor \isvarying{t} \lor \isvarying{f}
  \qquad
  \isvarying{\mathsf{EWhile}(c,b)} = \isvarying{c} \lor \isvarying{b}
\]
\[
  \isvarying{\mathsf{EFor}(lo,hi,b)} = \isvarying{lo} \lor \isvarying{hi} \lor \isvarying{b}
\]
\[
  \isvarying{\mathsf{ESeq}(\bar{e})} = \exists\, e_i \in \bar{e}.\;\isvarying{e_i}
  \qquad
  \isvarying{\mathsf{EApp}(\bar{e})} = \exists\, e_i \in \bar{e}.\;\isvarying{e_i}
\]
\[
  \isvarying{\mathsf{ELet}(x,v,b)} = \isvarying{v} \lor \isvarying{b}
  \qquad
  \isvarying{\mathsf{ESuperstep}(\_,body,cont)} = \isvarying{body} \lor \isvarying{cont}
\]

\subsection{$\chk{m}{e}$ — the barrier-error collector}

\[
  \chk{m}{\mathsf{ELit}} = []
  \qquad
  \chk{m}{\mathsf{EVary}} = []
  \qquad
  \chk{m}{\mathsf{EVar}(x)} = []
\]
\[
  \chk{m}{\mathsf{EBarrier}} =
  \begin{cases}
    [\mathsf{BarrierError}] & \text{if } m = \Diverged \\
    [] & \text{if } m = \Converged
  \end{cases}
\]
\[
  \chk{m}{\mathsf{EBinop}(a,b)} = \chk{m}{a} \mathbin{++} \chk{m}{b}
  \qquad
  \chk{m}{\mathsf{EUnop}(e)} = \chk{m}{e}
\]
\[
  \chk{m}{\mathsf{ELet}(x,v,b)} = \chk{m}{v} \mathbin{++} \chk{m}{b}
\]
Let $m' = \text{if } \isvarying{c} \text{ then } \Diverged \text{ else } m$.
\[
  \chk{m}{\mathsf{EIf}(c,t,f)} = \chk{m}{c} \mathbin{++} \chk{m'}{t} \mathbin{++} \chk{m'}{f}
\]
\[
  \chk{m}{\mathsf{EWhile}(c,b)} = \chk{m}{c} \mathbin{++} \chk{m'}{b}
  \qquad
  \chk{m}{\mathsf{EFor}(lo,hi,b)} = \chk{m}{lo} \mathbin{++} \chk{m}{hi} \mathbin{++} \chk{m''}{b}
\]
where $m'' = \text{if } \isvarying{lo} \lor \isvarying{hi} \text{ then } \Diverged \text{ else } m$.
\[
  \chk{m}{\mathsf{ESeq}(\bar{e})} = \mathbin{++}_{e_i \in \bar{e}} \chk{m}{e_i}
  \qquad
  \chk{m}{\mathsf{EApp}(\bar{e})} = \mathbin{++}_{e_i \in \bar{e}} \chk{m}{e_i}
\]

Let $\mathit{entry}(m, b) = [\mathsf{BarrierError}]$ if $m = \Diverged$ and
$b = \mathsf{false}$, and $[]$ otherwise.
\[
  \chk{m}{\mathsf{ESuperstep}(b,\mathit{body},\mathit{cont})}
  = \mathit{entry}(m,b) \mathbin{++} \chk{m}{\mathit{body}} \mathbin{++} \chk{m}{\mathit{cont}}
\]
A non-divergent superstep (\,$b = \mathsf{false}$\,) entered under $\Diverged$
mode raises $\mathsf{BarrierError}$ because the implicit end-of-superstep barrier
is reached by only some threads.

\subsection{$\cdcf{e}$}

\begin{align*}
  &\cdcf{\mathsf{EIf}(c,t,f)} = \isvarying{c} \lor \cdcf{t} \lor \cdcf{f} \\
  &\cdcf{\mathsf{EWhile}(c,b)} = \isvarying{c} \lor \cdcf{b} \\
  &\cdcf{\mathsf{EFor}(lo,hi,b)} = \isvarying{lo} \lor \isvarying{hi} \lor \cdcf{b} \\
  &\cdcf{\mathsf{ELet}(x,v,b)} = \cdcf{v} \lor \cdcf{b} \\
  &\cdcf{\mathsf{ESuperstep}(\_,body,cont)} = \cdcf{body} \lor \cdcf{cont} \\
  &\cdcf{e} = \mathsf{false} \quad \text{(all other constructors, incl.\ $\mathsf{EVar}$)}
\end{align*}

%───────────────────────────────────────────────────────────────────────────────
\section{Properties}
\label{sec:properties}

\subsection{T0 --- Dimension-usage semilattice}
\label{sec:dimmerge}

\begin{theorem}[Merge commutativity]
  \label{thm:merge-comm}
  For all $d_1, d_2$: $d_1 \mergedim d_2 = d_2 \mergedim d_1$.
\end{theorem}

\begin{theorem}[Merge associativity]
  \label{thm:merge-assoc}
  For all $d_1, d_2, d_3$: $(d_1 \mergedim d_2) \mergedim d_3 = d_1 \mergedim (d_2 \mergedim d_3)$.
\end{theorem}

\begin{theorem}[Merge idempotence]
  \label{thm:merge-idem}
  For all $d$: $d \mergedim d = d$.
\end{theorem}

\begin{theorem}[Right identity]
  \label{thm:merge-empty-r}
  For all $d$: $d \mergedim \emptydim = d$.
\end{theorem}

\begin{theorem}[Left identity]
  \label{thm:merge-empty-l}
  For all $d$: $\emptydim \mergedim d = d$.
\end{theorem}

\subsection{T0 --- Sequential composition homomorphism}

\begin{theorem}[Check-seq homomorphism]
  \label{thm:seqhom}
  For any mode $m$ and expression lists $\bar{e}_1, \bar{e}_2$:
  \[
    \chk{m}{\mathsf{ESeq}(\bar{e}_1 \mathbin{++} \bar{e}_2)}
    = \chk{m}{\mathsf{ESeq}(\bar{e}_1)} \mathbin{++} \chk{m}{\mathsf{ESeq}(\bar{e}_2)}.
  \]
\end{theorem}

\subsection{T1 --- Barrier-safety invariants}

\begin{theorem}[Diverged-clean iff barrier-free]
  \label{thm:div-clean}
  For all $e$:
  \[
    \chk{\Diverged}{e} = []
    \;\iff\;
    \barrierfree{e} = \mathsf{true}.
  \]
\end{theorem}

\begin{theorem}[Mode monotonicity]
  \label{thm:monotone}
  For all $e$:
  \[
    \chk{\Converged}{e} \incl \chk{\Diverged}{e}.
  \]
  (Strengthening the execution context never removes errors.)
\end{theorem}

\begin{lemma}[Non-varying implies clean in converged mode]
  \label{thm:nonvar-clean}
  For all $e$:
  \[
    \isvarying{e} = \mathsf{false}
    \;\Rightarrow\;
    \chk{\Converged}{e} = [].
  \]
\end{lemma}

\begin{theorem}[CDCF agreement]
  \label{thm:cdcf-clean}
  For all $e$:
  \[
    \cdcf{e} = \mathsf{false}
    \;\Rightarrow\;
    \chk{\Converged}{e} = [].
  \]
\end{theorem}

\begin{theorem}[Varying condition flags barriers]
  \label{thm:varying-flags}
  For all $e_c, e_t, e_f$:
  \[
    \isvarying{e_c} = \mathsf{true}
    \;\land\;
    \barrierfree{e_t} = \mathsf{false}
    \;\Rightarrow\;
    \chk{\Converged}{\mathsf{EIf}(e_c, e_t, e_f)} \ne [].
  \]
\end{theorem}

\begin{theorem}[Outer-diverged superstep raises error (F-01)]
  \label{thm:superstep-outer-diverged}
  For all $\mathit{body}, \mathit{cont}$:
  \[
    \chk{\Diverged}{\mathsf{ESuperstep}(\mathsf{false},\,\mathit{body},\,\mathit{cont})} \ne [].
  \]
  A non-divergent superstep entered under a $\Diverged$ execution context
  unconditionally produces a $\mathsf{BarrierError}$.
\end{theorem}

\subsection{T1 --- Env-threaded model (T2-F02)}
\label{sec:envthreaded-props}

\begin{theorem}[Env-lookup after env-extend (same key)]
  \label{thm:env-lookup-extend-same}
  For all $\mathit{env}$, $x$, $v$:
  \[
    \mathtt{env\_lookup}(\mathtt{env\_extend}(\mathit{env},\,x,\,v),\;x) = v.
  \]
\end{theorem}

\begin{theorem}[Let-alias variability propagation]
  \label{thm:env-let-alias-varying}
  For all $\mathit{env}$, $x$, $v$, $b$, if
  $\mathtt{is\_varying\_in\_env}(\mathit{env}, v) = \mathsf{true}$ and
  $\mathtt{is\_varying\_in\_env}(\mathtt{env\_extend}(\mathit{env},x,\mathsf{true}), b) = \mathsf{true}$,
  then
  \[
    \mathtt{is\_varying\_in\_env}(\mathit{env},\;\mathsf{ELet}(x,v,b)) = \mathsf{true}.
  \]
\end{theorem}

\begin{theorem}[EVar carries no barrier (F-02 auxiliary)]
  \label{thm:env-var-diverged-clean}
  For all $\mathit{env}$, $x$:
  \[
    \mathtt{check\_env}(\Diverged,\,\mathit{env},\,\mathsf{EVar}(x)) = [].
  \]
\end{theorem}

\begin{theorem}[Let-alias soundness (F-02)]
  \label{thm:env-check-let-alias-catches}
  For all $\mathit{env}$, $x$, $v$, if
  $\mathtt{is\_varying\_in\_env}(\mathit{env}, v) = \mathsf{true}$, then
  \[
    \mathtt{check\_env}(\Converged,\,\mathit{env},\;
      \mathsf{ELet}(x,v,\mathsf{EIf}(\mathsf{EVar}(x),\mathsf{EBarrier},\mathsf{ELit})))
    \ne [].
  \]
  A barrier guarded by a let-aliased varying variable is flagged even in
  $\Converged$ mode.
\end{theorem}

%───────────────────────────────────────────────────────────────────────────────
\section{Env-threaded model}
\label{sec:envthreaded}

The binding-blind functions $\isvarying{e}$ and $\chk{m}{e}$ treat every
$\mathsf{EVar}(x)$ as non-varying regardless of what was bound to $x$.
This mirrors the pre-fix \texttt{Sarek\_convergence.ml} behaviour.
Finding F-02 showed that let-aliased thread-varying values were therefore
missed; the OCaml fix propagates variability through \texttt{TELet}.

The env-threaded model formalises the corrected behaviour.

\begin{definition}[Environment]
\label{def:env}
$\mathsf{Env} = \mathsf{list}(\mathbb{N} \times \mathbb{B})$
is an association list mapping variable identifiers to their variability flags.
\[
  \mathtt{env\_lookup}(\mathit{env},\,x) =
  \begin{cases}
    v & \text{if } (x,v) \text{ is the first pair in } \mathit{env} \text{ with key } x \\
    \mathsf{false} & \text{if no such pair exists}
  \end{cases}
\]
\[
  \mathtt{env\_extend}(\mathit{env},\,x,\,v) = (x,v) :: \mathit{env}
\]
\end{definition}

\begin{definition}[$\mathtt{is\_varying\_in\_env}(\mathit{env}, e)$]
\label{def:isvarying-env}
Env-threaded variability; extends $\isvarying{\cdot}$ with:
\[
  \mathtt{is\_varying\_in\_env}(\mathit{env},\,\mathsf{EVar}(x))
  = \mathtt{env\_lookup}(\mathit{env},\,x)
\]
\[
  \mathtt{is\_varying\_in\_env}(\mathit{env},\,\mathsf{ELet}(x,v,b))
  = \mathtt{is\_varying\_in\_env}(
      \mathtt{env\_extend}(\mathit{env},\,x,\,\mathtt{is\_varying\_in\_env}(\mathit{env},v)),\;b)
\]
All other constructors recurse with the same $\mathit{env}$ (unchanged).
\end{definition}

\begin{definition}[$\mathtt{check\_env}(m, \mathit{env}, e)$]
\label{def:checkenv}
Env-threaded barrier-error collector; extends $\chk{m}{\cdot}$ with:
\[
  \mathtt{check\_env}(m,\,\mathit{env},\,\mathsf{EVar}(x)) = []
\]
\[
  \mathtt{check\_env}(m,\,\mathit{env},\,\mathsf{ELet}(x,v,b))
  = \mathtt{check\_env}(m,\,\mathit{env},\,v) \mathbin{++}
    \mathtt{check\_env}(m,\,\mathtt{env\_extend}(\mathit{env},x,\mathtt{is\_varying\_in\_env}(\mathit{env},v)),\,b)
\]
For $\mathsf{EIf}$/$\mathsf{EWhile}$/$\mathsf{EFor}$ the inner mode is
determined by $\mathtt{is\_varying\_in\_env}$ instead of $\isvarying{\cdot}$.
All other constructors are structurally identical to $\chk{m}{\cdot}$.
\end{definition}

%───────────────────────────────────────────────────────────────────────────────
\section{Refinement to \texttt{sarek/ppx/}}
\label{sec:refinement}

\subsection*{Type correspondence}

The abstract \texttt{expr} type maps to the real \texttt{texpr} AST as follows
(see \texttt{ASSUMPTIONS.md} for the full table and elided constructors):

\begin{center}
\begin{tabular}{l l l}
\hline
Abstract & Real & Notes \\
\hline
\texttt{EVary} & \texttt{TEVar} (varying), \texttt{TEIntrinsicConst} (varying) & \texttt{is\_thread\_varying} recurses \\
\texttt{EVar}($x$) & \texttt{TEVar} (by id) & variability resolved via env (T2-F02) \\
\texttt{EBarrier} & \texttt{TEIntrinsicFun \_ (Some ConvergencePoint) \_} & error when mode=Diverged \\
\texttt{ELit} & all uniform literals + \texttt{TEUnit}/\texttt{TEBool}/\texttt{TEInt}/... & \\
\texttt{EIf} & \texttt{TEIf}, \texttt{TEMatch} & \\
\texttt{EFor} & \texttt{TEFor} & lo/hi checked for variation \\
\texttt{ESeq} & \texttt{TESeq} & \\
\texttt{ELet}($x$,$v$,$b$) & \texttt{TELet}, \texttt{TELetMut} & env-threaded model propagates variability (T2-F02) \\
\texttt{ESuperstep}($b$,$body$,$cont$) & \texttt{TESuperstep} & F-01: diverged-entry BarrierError (Phase 1a) \\
\texttt{EApp} & \texttt{TEApp}, \texttt{TETuple}, \texttt{TERecord} & \\
\texttt{EBinop} & \texttt{TEBinop}, \texttt{TEVecSet}, \texttt{TEArrSet}, \texttt{TEAssign} & \\
\texttt{EUnop} & \texttt{TEUnop}, \texttt{TEFieldGet}, \texttt{TEReturn}, \texttt{TECreateArray} & \\
\hline
\end{tabular}
\end{center}

\subsection*{Conformance claim}

For all \texttt{texpr} inputs $t$ that map to an abstract expression
$e = \mathtt{texpr\_to\_expr}(t)$ (i.e.\ $t$ does not contain any elided
constructor), and for each mode $m$:
\[
  \mathtt{normalize}(\mathtt{real\_check\_expr}(\{mode=m\}, t))
  = \chk{m}{e}
\]
where $\mathtt{normalize}$ strips metadata from \texttt{texpr}-level errors and
projects to the abstract \texttt{error} type.

\subsection*{Elided constructors (out of scope)}

\texttt{TELetShared}, \texttt{TELetRec},
\texttt{TEPragma}, \texttt{TEOpen}, \texttt{TENative}, \texttt{TEGlobalRef},
and the \texttt{WarpConvergence} error class.
(\texttt{TESuperstep} is now in scope as $\mathsf{ESuperstep}$; see type-correspondence table.)

%───────────────────────────────────────────────────────────────────────────────
\section{References}
\label{sec:references}

\begin{itemize}
  \item SPOC source: \url{https://github.com/mathiasbourgoin/SPOC}
  \item \texttt{sarek/ppx/Sarek\_convergence.ml} --- target implementation
  \item \texttt{ASSUMPTIONS.md} --- elided-constructor rationale
  \item Sarek formal apparatus memory: \texttt{~/.claude/projects/.../memory/sarek-convergence-formal.md}
\end{itemize}

\end{document}
